On the Functionality of the Multilayer Perceptron

### 2020-09-05 Thesis/Sections/TheNerualNetwork.tex

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage[top=2.5cm,left=2.5cm,right=2.5cm]{geometry}

@@

\paragraph*{}In the present age, some problems have shown themselves to be increasingly difficult to solve through conventional computer programs. The challenge arose as to how to build some sort of computer program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for an algorithm, we seek an architecture that allows for a computer to \textit{learn} and change and update itself as it is presented with more data. Such an algorithm exists in the form of a \textit{Neural Network} \cite{Geron2,Goodfellow,Levine}.

\paragraph*{}A neural network is a mathematical function that seeks to produce an approximation $F^*$, of some unknown function $F$. For example, in a $k$-classes classifications setting, a network model is provided with some input array $x$ and tasked to map it to a particular categorical output $0$ through $k-1$, using a set of parameters, $\Theta$ \cite{Goodfellow,James}. During a period of training, a model is presented with sample inputs $x$, an expected output $y$, and optimizes the parameters of the set $\Theta$ in order to produce the approximated prediction, $y^*$ which maps to one of $k$ outputs.

\paragraph*{}Former YouTube video Classification team lead, and current Machine Learning consultant, Aurelien Geron writes \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result of such an analogy is a computer program that is reminiscent of the brain. Typically, a computer program is simply a set of mathematical or logical instructions given to a computer to execute. In 1943, Neuroscientist Warren McCulloch and Mathematician Walter Pitts published a paper \cite{McCulloch} which began to pave the way for developing such a mathematical model of the brain \cite{Geron,Goodfellow,Loy}. This model has been adapted over the last few decades to be able to produce an model that can functionally teach itself how to solve a very particular problem.

@@

\paragraph*{}A Neural Network is simply a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Geron,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. It is this process of transforming inputs successively in a particular order until an output is attained \cite{Geron,Loy}. This output encodes the models final "decision" given a unique input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are name \textit{networks} because of their nested composition nature. The model can be represented by a linear or acyclic graphs which successively maps exactly how the repeated composition is structured. In a \textit{feed-forward network}, information is passes successively in one direction. For example functions could be chained together such as \cite{Goodfellow}
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
Where each function $f^{(l)}$ represents a layer of the network model. The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.

@@

\paragraph*{}Typically, a layer function can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. Many feed-forward network layers follow this structure. Step (i.) is usually in the form of a matrix-vector equation:
\begin{equation}
\label{eqn-matVectEqn}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
Where $W^{(l)}$ is the \textit{weighting-matrix} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.
Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-matVectEqn}). $\sigma^{(l)}$ is some activation function, which is often use to enable the modeling of more complex-decision boundaries.

Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

@@

@@

@@

>>>>

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[top=2.5cm,left=2.5cm,right=2.5cm]{geometry}

@@

\paragraph*{}The task of algorithmically matching sound files to musical instrument classes in exceedingly difficult through conventional computer programming techniques. Because of the complexity of soundwaves, and human ears and brains interpret them, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an architecture that allows for a computer to \textit{learn} and change and update itself as it is presented with more data. Such an algorithm exists in the form of a \textit{Neural Network} \cite{Geron2,Goodfellow,Levine}.

\paragraph*{}A neural network is a mathematical function that seeks to produce an approximation $F^*$, of some usually unknown function $F$ . For audio recognition, $F$ is some function, or series of functions, that occurs in the brain which allows for a listener to take the sensation of audio, and map it to a known musical instrument. For a neural network, a model must construct an approximate function $F^*$ using a set of parameters $\Theta$, which allows the model to map a series of inputs, $x$ (called \textit{features}) to a series of outputs, $y$ (called \textit{predictions}) \cite{Goodfellow,James,Virtanen}.

\paragraph*{}Former YouTube video Classification team lead, and current Machine Learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result of such an analogy is a computer program that is reminiscent of the brain.

\paragraph*{}Over the course of their lives, humans will learn to map sounds to sources almost effortlessly. Given a sample of sound, and the appropriate label, humans can do this with very reasonable accuracy over a wide array of sounds \cite{Olsen,White}. Similarly, a neural network uses features of a sound, represented by an array $x$, and learns a series of parameters, $\Theta$, that allow it to map those features to one or more pre-defined sources \cite{Virtanen}.

\paragraph*{}For this particular project, a neural network has been constructed to perform a  \textit{classification} task. A classification task involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument that could have produced the sound-wave in the audio file.

@@

\paragraph*{}A Neural Network is simply a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. It is this process of transforming inputs successively in a particular order until an output is attained \cite{Geron,Loy}. This output encodes the models final "decision" given a unique input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are name \textit{networks} because of their nested composition nature. The model can be represented by a linear or acyclic computational graph which successively maps exactly how the repeated composition is structured. In a \textit{feed-forward network}, information is passes successively in one direction. For example functions could be chained together such as \cite{Goodfellow}
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
Where each function $f^{(l)}$ represents a layer of the network model. The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.

@@

\paragraph*{}Typically, a layer function can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries. Step (i.) is usually in the form of a matrix-vector equation:
\begin{equation}
\label{eqn-matVectEqn}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
Where $W^{(l)}$ is the \textit{weighting-matrix} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.
Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-matVectEqn}). $\sigma^{(l)}$ is some activation function, which is often use to enable the modeling of more complex-decision boundaries.

\paragraph*{}Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

% ================

@@

% ================

@@

\paragraph*{}The success of a neural network algorithm is enormously dependent of the strength of the chosen features, rather than the quantity of available data \cite{Geron,Khan,Liu}. We have derived an appropriate set of inputs arrays \textit{[See Section on Features]}, and constructed a neural network architecture to compliment these features.

\paragraph*{}The full classification neural network used for this project consists of two distinct entry points. This means that rather than presenting the network with one set of input data, $X$, we present the network with two different arrays, $X_1$ and $X_2$. Both arrays are a produce of the same audio file sound wave, and thus share a common training label. Each of the two entry points feed into their own, non-related set of layer functions. The second-to-last layer, the outputs from both neural networks are concatenated into a single dense layer, which is then passed into the output layer, which contains $k$ neurons, one of reach unique class \cite{Loy,Geron}. A visual representation of this architecture can be found in fig. (\ref{fig-NetworkArchitecture}).

\begin{figure}[H]
\label{fig-NetworkArchitecture}
\begin{center}
\includegraphics[scale=0.45]{../Figures/NeuralNetworkArchitecture}
\end{center}
\caption{The developed architecture of the audio file classification neural network. The Left branch process an image-like input, the right branch processes a vector-like input. The activations are then merged, and then a single output is produced}
\end{figure}

% ================

\subsubsection{The Spectrogram Branch}

\paragraph*{}The spectrogram branch is pictured on the left side of fig. (\ref{fig-NetworkArchitecture}). A spectrogram is a representation of a sound wave, or signal by representing energy distributions as a function of \textit{time} and \textit{frequency} \cite{White,Olsen,Kahn}. \textcolor{blue}{Details on the creating of the spectrogram can be found in the \textit{features} section of this document.}

\paragraph*{}The input layer of the spectrogram accepts a 4-Dimensional array. The axes, in order of indexing, represents (i) the size of the \textit{mini-batch} of samples, (ii) the pixel width of each sample, (iii) the pixel height of each sample, and (iv) the number of channels in each sample. As a model hyper-parameter, we have chosen each batch to contain $64$ samples. We have also chosen to truncate time and frequency axes (ii \& iii) to contain $560$ frequency bins, and $256$ time-frames. Each image also contains a single channel, which make it gray-scale when visualized. We can denote the 4D shape of the input object into this branch as:
\begin{equation}
\label{eqn-shapeX1}
X_1 \in \mathbb{R}^{(64 \times 560 \times 256 \times 1)}
\end{equation}
Any other shape will be rejected by the model, and an error is raised.

\paragraph*{}After the input layer assert the appropriate shape, the array $X_1$, which is a collection of $64$ spectrograms, is passed into the first of three \textit{Convolution Layer Groups}. These layer groups are inspired from the \textit{VGG-16 Neural Network} architecture \textcolor{red}{[citation needed]}. Each convolution layer group in composed of three individual layers: 
\begin{enumerate}
\item A 2-Dimensional Convolution layer, using $32$ filters, a $3 \times 3$ kernel, a $1 \times 1$ step size, and a ReLU activation function,
\item A 2-Dimensional Convolution layer, using $32$ filters, a $3 \times 3$ kernel, a $1 \times 1$ step size, and a ReLU activation function,
\item A 2-Dimensional Maximum Pooling layer using a $3 \times 3$ pooling size, and an Identity activation function
\end{enumerate}  
The Convolution layers convoluted over the middle two axes of the data (over space and time in the spectrogram).

\paragraph*{}By grouping layers in this structure, each of the $32$ filters passes over a $3 \times 3$ area, and then again. The result is

% ================

\subsubsection{The Perceptron Branch}

\paragraph*{} The Perceptron branch is picture on the right side of fig. (\ref{fig-NetworkArchitecture}), notice that it is considerably smaller in side and complexity than it's neighbor. Rather than accept an image-like input, the perception simply takes a vector-like input of properties of a larger data set - in this case, each waveform. We call these properties \textit{features} \cite{Geron,Kahn,Virtanen}. \textcolor{blue}{Details on the creating of these features can be found in the \textit{features} section of this document.}

\paragraph*{}The input layer of the perceptron accepts a 2-Dimensional array. The axes, in order of indexing, represent (i) the size of the \textit{mini-batch} of samples, (ii) the number of features for each sample. We use the same model hyper-paramater of $64$ samples per batch, and have developed $23$ unique clssification features that have been derived from time-space and frequency-space representations of the audio file data. We can denote the 2D shape if the input object into this branch as:
\begin{equation}
\label{eqn-shapeX2}
X_2 \in \mathbb{R}^{(64 \times 23)}
\end{equation}
Again, any other shape will be rejected by the model, and an error is raised.

% ================

\subsubsection{The Final Output Branch}

\paragraph*{}

@@

\bibitem{Olsen}
Olson, Harry E. \textit{Music, Physics and Engineering}. 2nd ed., Dover Publications, 1967.
